Environmental factors affecting Galaxy Morphology

### 2014-07-27 Writeup/paper_v2.tex
comments through page 3

\author[Kannawadi et al.]
{Arun Kannawadi$^1$\thanks{\tt akannawa@andrew.cmu.edu}, 
Rachel Mandelbaum$^1$,
Claire Lackner$^2$, 
\\$^1$McWilliams Center for Cosmology, Carnegie Mellon University, Pittsburgh, PA 15217, USA
\\$^2$Kavli Institute for the Physics and Mathematics of the Universe (WPI), Todai Institutes for Advanced Study, the University of Tokyo, Kashiwa, Japan.
}

@@

\begin{abstract}
According to our current understanding, galaxy shapes and morphologies should depend on various factors such as the local environment. Realistic image simulations for calibration of weak lensing analysis methods that use training samples from the Hubble Space Telescope can therefore be affected by these trends, due to the limited volume of the universe that has been surveyed by Hubble. We will show how redshift slices in a volume-limited subsample of COSMOS can be classified as overdense or underdense (or neither), and how the statistical properties of various morphological parameters such as ellipticity, \sersicn, \btt ratio and color differ in these bins. This study requires a careful distinction between environment effects from large-scale structure, which we do not wish to include in simulations, and general trends in the galaxy population with redshift. We conclude with some guidance for how upcoming surveys can use COSMOS data as the basis for weak lensing simulations without having their conclusions overly affected by cosmic variance.  
\end{abstract}

@@

\section{Introduction}
\label{S:intro}
{\bf \emph{Stress on the importance of WL:}} We are currently in the era of precision cosmology, where systematic errors typically dominate over statistical errors. In almost all branches of observational cosmology,
weak lensing introduces systematic error, in addition to itself being a probe of the dark matter and eventually the large scale structure of the universe.
As light from the background sources passes through the foreground cluster of galaxies and dark matter profiles, the shapes get distorted. Weak lensing also causes magnification in addition to introducing a shear.
Weak lensing effects need to be removed from the Cosmic Microwave Background (CMB) analysis to understand the initial conditions of the universe.

@@

%The structure of the paper is as follows. In Sec.~\ref{S:data}, we briefly describe the COSMOS catalog that we use for our analysis.
In this paper, we show that these large scale structure effects mentioned are indeed present in the data. The rest of the paper is structured as follows.
In Sec.~\ref{sub:overdensities}, we explain how we volume-limit our data and identify overdensities and underdensities and in Sec.~\ref{sub:axisratio},
 we explain how we obtain axis-ratios from the images of the galaxies. Finally, we present our results in Sec.~\ref{S:results} and end with conclusions in Sec.~\ref{S:conclusion}.
\section{Data}
\label{S:data}
%In this analysis, we consider the COSMOS sample of galaxies. 
COSMOS~\citep{COSMOS_overview, COSMOS_generic, COSMOS_Alexie} is a flux-limited, narrow deep field survey covering a contiguous area of $1.64 \text{ deg}^2$ of sky, with images taken using the Advanced Camera for Surveys (ACS) Wide Field Channel (WFC)
in the Hubble Space Telescope (HST). Precise shape measurements, when compared to ground-based surveys, can be made since the full width half-maximum (FWHM) of the point-spread function (PSF)
is $0.12''$. High resolution images taken through the wide F814W filter (broad \emph{I}), after correcting for the PSF, have led to collection of postage stamp images\footnote{http://irsa.ipac.caltech.edu/data/COSMOS/images/galaxy-postage-stamps/}, originally intended for simulations of ground-based data using SHERA (SHEar Reconvolution Analysis)~\citep{SHERA}.
\cite{Claire_Fits} fitted parametric models to most of these galaxies including \sersicn profile fits, 
2 component bulge+disk fits, axis ratios etc. In addition to the ACS/WFC (F814W) imaging, the COSMOS field has also been imaged by Subaru Suprime-Cam ($B_j, V_j, g^+,r^+,i^+,z^+,$ NB816), the
Canada-French Hawaii Telescope (CFHT;$u^*,i^*$) and the KPNO/CTIO ($K_s$) .

Photometric redshifts were determined by~\cite{COSMOS_photoz}.The accuracy of photometric redshifts for F814W $\le$ 22.5 is $\sigma_{\Delta z} = 0.031(1+z_s)$.
The photometric redshift values become noisy beyond $z$ of $1$ for our purposes and the various fits to the galaxies are also not very reliable beyond the apparent magnitude $(m)$ value of 23.5. ($S/N \sim 50$ {\bf ?} )

@@

\section{Methods}
\label{S:methods}
\subsection{Finding overdensities}
\label{sub:overdensities}
Every survey is flux-limited and so is COSMOS, and is affected by Malmquist bias.
We generate, using the method below, a volume-limited sample that is complete upto $z=1$ by applying a cut on luminosity such that only galaxies instrinsically brighter than a certain threshold is considered. This threshold is set on $K$-corrected $i$-band magnitudes, denoted as $M_I$.
Since the parent sample contains fainter galaxies, upto F814W=25.2, we compare the distribution of the F814W=23.5 sample with the samples containing fainter galaxies for high redshift bins, to see where the sample is no longer complete.
\begin{figure}
 \centering
 a) \includegraphics[width=\columnwidth]{volume_limiting_pdf}
 b) \includegraphics[width=\columnwidth]{volume_limiting_cdf}
 \label{fig:volume_limiting_dist}
 \caption{ ... }
\end{figure}

@@

In a wide-field survey, regions of overdensities are identified by computing 3-dimensional comoving densities. But in narrow field surveys like COSMOS, previous work {\bf cite } have shown
that regions of overdensities can be identified from a 1-dimensional histogram of redshifts, at least for low redshifts.

For our sample of galaxies, volume-limited upto $z=1$ by imposing an $M_I$ cut, we fit parametric models to the histogram of photometric redshifts in order to assign values of overdensity. The figures show chi-squared and gamma functions fitted to the histogram~\citep{Redshift_modelling}.
For low $z$, the volume is too small to rely the overdensity values from our model fits and hence the fit is made for $z \ge 0.3$. Our binning width, $\Delta z$ must be small enough so as to be able to identify localized overdensities/underdensities but large enough to not let our conclusions be affected by the errors in photometric redshift. 
We choose our bins to be $0.05$ wide starting from $z=0.3$.

>>>>

\author[Kannawadi et al.]
{Arun Kannawadi$^1$\thanks{\tt akannawa@andrew.cmu.edu}, 
Rachel Mandelbaum$^1$,
Claire Lackner$^2$, 
\\$^1$McWilliams Center for Cosmology, Carnegie Mellon University, Pittsburgh, PA 15217, USA
\\$^2$Kavli Institute for the Physics and Mathematics of the Universe (WPI), Todai Institutes for Advanced Study,\\ the University of Tokyo, Kashiwa, Japan.
}

@@

\begin{abstract}
According to our current understanding, galaxy shapes and morphologies should depend on various factors such as the local environment. Realistic image simulations for calibration of weak lensing analysis methods that use training samples from the Hubble Space Telescope can therefore be affected by these trends, due to the limited volume of the universe that has been surveyed by Hubble. We will show how redshift slices in a volume-limited subsample of COSMOS can be classified as overdense or underdense (or neither), and how the statistical properties of various morphological parameters such as ellipticity, \sersicn, \btt ratio and color differ in these bins. This study requires a careful distinction between environment effects from large-scale structure, which we do not wish to include in simulations, and general trends in the galaxy population with redshift. We conclude with some guidance for how upcoming surveys can use COSMOS data as the basis for weak lensing simulations without having their conclusions overly affected by cosmic variance.  
\rachel{I have some comments on this abtract, but prefer to do all
  revision of the abstract at the end once the paper is finalized.}
\end{abstract}

@@

\section{Introduction}
\label{S:intro}

\rachel{It's true that you should stress the importance of WL.  But
  I recommend restructuring this opening to be {\em constructive}
  rather than the current tone which is relatively negative.  To make
  it constructive, I suggest covering the following points in the
  following order:}
\begin{itemize}
\item Major surveys are planned in the next 1-2 decades
  to make measurements that will precisely constrain the nature of
  dark energy.
\item Weak lensing is one of the more important cosmological probes
  for this purpose.
\item For these upcoming surveys to achieve their promise, the
  systematic error budget must be below the statistical error budget.
  This means reducing systematic errors from their current typical
  level.
\item One method that is commonly used to test for the presence of
  systematics is image simulation.  Typically the most realistic ones
  use HST-based training samples based on COSMOS or the UDF, which are
  small areas of the sky.
\item THe goal of this work is to understand systematic uncertainties
  that might arise due to the finite size of these areas, when
  dividing the sample into redshift slices to test for shear
  systematics as a function of redshift.
\end{itemize}

\rachel{And then you can launch into some of your current text about why there
is cause for concern (known environment-morphology relations, etc.).}

{\bf \emph{Stress on the importance of WL:}} We are currently in the era of precision cosmology, where systematic errors typically dominate over statistical errors. In almost all branches of observational cosmology,
weak lensing introduces systematic error, in addition to itself being a probe of the dark matter and eventually the large scale structure of the universe.
As light from the background sources passes through the foreground cluster of galaxies and dark matter profiles, the shapes get distorted. Weak lensing also causes magnification in addition to introducing a shear.
Weak lensing effects need to be removed from the Cosmic Microwave Background (CMB) analysis to understand the initial conditions of the universe.

@@

%The structure of the paper is as follows. In Sec.~\ref{S:data}, we briefly describe the COSMOS catalog that we use for our analysis.
In this paper, we show that these large scale structure effects mentioned are indeed present in the data. The rest of the paper is structured as follows.
In Sec.~\ref{sub:overdensities}, we explain how we volume-limit our data and identify overdensities and underdensities and in Sec.~\ref{sub:axisratio},
 we explain how we obtain axis-ratios from the images of the galaxies. Finally, we present our results in Sec.~\ref{S:results} and end with conclusions in Sec.~\ref{S:conclusion}.
\section{Data}
\label{S:data}
%In this analysis, we consider the COSMOS sample of galaxies. 
COSMOS~\citep{COSMOS_overview, COSMOS_generic, COSMOS_Alexie} is a flux-limited, narrow deep field survey covering a contiguous area of $1.64 \text{ deg}^2$ of sky, with images taken using the Advanced Camera for Surveys (ACS) Wide Field Channel (WFC)
in the Hubble Space Telescope (HST). Precise shape measurements, when compared to ground-based surveys, can be made since the full width half-maximum (FWHM) of the point-spread function (PSF)
is $0.12''$. High resolution images taken through the wide F814W filter (broad \emph{I}), after correcting for the PSF, have led to collection of postage stamp images\footnote{http://irsa.ipac.caltech.edu/data/COSMOS/images/galaxy-postage-stamps/}, originally intended for simulations of ground-based data using SHERA (SHEar Reconvolution Analysis)~\citep{SHERA}.
\cite{Claire_Fits} fitted parametric models to most of these galaxies including \sersicn profile fits, 
2 component bulge+disk fits, axis ratios etc. In addition to the ACS/WFC (F814W) imaging, the COSMOS field has also been imaged by Subaru Suprime-Cam ($B_j, V_j, g^+,r^+,i^+,z^+,$ NB816), the
Canada-French Hawaii Telescope (CFHT;$u^*,i^*$) and the KPNO/CTIO ($K_s$) .
\rachel{Technically speaking, \cite{Claire_Fits} did {\em not} fit
  models to these data.  Claire fit models to these data using the
  methodology of \cite{Claire_Fits}, and the appropriate reference
  describing these specific fits is the GREAT3 handbook.  However,
  given that you use them, you do need to say a bit more about them
  then you have already done.  State clearly what models were fit,
  with what free parameters. Finally, be clear about which sample you
  are using.  The SHERA paper only went to 22.5, and the GREAT3
  challenge used galaxies to 23.5.  Since you're using the latter
  sample, you should be very clear about it; currently it seems
  ambiguous. Should also mention the presence of the $I<25.2$ parent
  sample, since you use it quite a bit. (Also, on a more minor note, it is not
  common to call them \sersicn{} profiles, just \sersic{} profiles.)}

Photometric redshifts were determined by~\cite{COSMOS_photoz}.The accuracy of photometric redshifts for F814W $\le$ 22.5 is $\sigma_{\Delta z} = 0.031(1+z_s)$.
The photometric redshift values become noisy beyond $z$ of $1$ for our purposes and the various fits to the galaxies are also not very reliable beyond the apparent magnitude $(m)$ value of 23.5. ($S/N \sim 50$ {\bf ?} )
\rachel{Our catalog has the more up-to-date Ilbert et al. photo-$z$,
  not the older Mobasher ones.  Need to fix citations. Also, we care
  about the photo-$z$ accuracy beyond 22.5, so it's important to give
  some information about it to the depth that we actually use.  Finally, I
  recommend against mentioning the $S/N$ because the calculation is
  complicated by the correlated noise, so most people use use
  magnitude instead of $S/N$.}

@@

\rachel{If we decide to use stellar masses, then we need to describe
  the stellar mass estimates.}

\section{Methods}
\label{S:methods}

\rachel{Here you should put a few introductory sentences about the
  purpose of this section. Could be something like ``To study the
  variation in the intrinsic ellipticity distribution with
  environment, there are two main steps.  The first is to identify
  redshift ranges with overdense and underdense environments.  The
  second is to estimate galaxy axis ratios.  In this section we
  describe how these two steps were done.''}

\subsection{Finding overdensities}
\label{sub:overdensities}

Every survey is flux-limited and so is COSMOS, and is affected by
Malmquist bias. \rachel{Not all surveys are flux-limited.}
\rachel{Explain why you are doing this.  Before saying {\em how} you
  generate a volume-limited sample, explain {\em why} you are
  generating one.}
We generate, using the method below, a volume-limited sample that is complete upto $z=1$ by applying a cut on luminosity such that only galaxies instrinsically brighter than a certain threshold is considered. This threshold is set on $K$-corrected $i$-band magnitudes, denoted as $M_I$.
Since the parent sample contains fainter galaxies, upto F814W=25.2, we compare the distribution of the F814W=23.5 sample with the samples containing fainter galaxies for high redshift bins, to see where the sample is no longer complete.
\begin{figure}
 \centering
 a) \includegraphics[width=\columnwidth]{volume_limiting_pdf}
 b) \includegraphics[width=\columnwidth]{volume_limiting_cdf}
 \label{fig:volume_limiting_dist}
 \caption{ ... }
\end{figure}

@@

In a wide-field survey, regions of overdensities are identified by computing 3-dimensional comoving densities. But in narrow field surveys like COSMOS, previous work {\bf cite } have shown
that regions of overdensities can be identified from a 1-dimensional histogram of redshifts, at least for low redshifts. 
\rachel{This is not entirely the point.  As we've discussed, the point
  is that we do not care about 3d densities for this particular science.  Weak lensing simulations
  will use galaxies selected in 2d slices, and hence we will simply do
  it that way as well.  But we use the comparison with a 3d analysis
  as a test of our method.}

\rachel{I thought we were only going up to $z=0.85$ but here you say $1$?}
For our sample of galaxies, volume-limited upto $z=1$ by imposing an $M_I$ cut, we fit parametric models to the histogram of photometric redshifts in order to assign values of overdensity. The figures show chi-squared and gamma functions fitted to the histogram~\citep{Redshift_modelling}.
For low $z$, the volume is too small to rely the overdensity values from our model fits and hence the fit is made for $z \ge 0.3$. Our binning width, $\Delta z$ must be small enough so as to be able to identify localized overdensities/underdensities but large enough to not let our conclusions be affected by the errors in photometric redshift. 
We choose our bins to be $0.05$ wide starting from $z=0.3$. 
\rachel{Were you going to use other fitting functions as a test of the
  method?  If so, you should introduce them here.}
